---
title: "Reproducible Research - Project1"
author: "Alexey Bolfaine"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(broom)
opts_chunk$set(echo = TRUE, results = 'hold')
```

## 1 - GETTING & LOAD THE DATA

```{r}
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', 'dataset.zip', method = 'curl')
unzip('dataset.zip')
```

```{r}
activity <- read.csv("activity.csv", header = TRUE)
```

## 2 - HISTOGRAM - TOTAL NUMBER OF STEPS PER DAY

```{r}
StepsPerDay <- tapply(activity$steps, activity$date, sum)
```

```{r}
hist(StepsPerDay, xlab = "Number of Steps", main = "Histogram - Steps per Day")
```

## 3 - CALCULATE MEAN AND MEDIAN OF STEPs TAKEN PER DAY

```{r}
MeanPerDay <- mean(StepsPerDay, na.rm = TRUE)
MedianPerDay <- median(StepsPerDay, na.rm = TRUE)
```
## 4 - PLOT AVERAGE NUMBER OF STEPS TAKEN

```{r}
StepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(StepsPerInterval)), 
     StepsPerInterval, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Average Daily Activity Pattern", 
     type = "l")
```

##  5 - The 5-minute interval that, on average, contains the maximum number of steps

```{r}
maxInterval <- names(sort(StepsPerInterval, decreasing = TRUE)[1])
maxSteps <- sort(StepsPerInterval, decreasing = TRUE)[1]
```
The 5-minute interval that, on average, contains the maximum number of steps is interval `r maxInterval`, at `r maxSteps.

## 6 - Code to describe and show a strategy for imputing missing data

```{r}
NA.vals <- sum(is.na(activity$steps))
```
Number of missing values: `r NA.vals`

## Here we fill the missing data with avg number of step, replace missing by avg number of steps taken.

```{r}
MissingData <- activity[is.na(activity$steps),]
DataWithoutMissing <- activity[!is.na(activity$steps),]
Mean_without_missing <- with(DataWithoutMissing, tapply(steps, DataWithoutMissing$interval, mean))
MissingData$steps <- Mean_without_missing
MergeDataActivity <- rbind(DataWithoutMissing, MissingData)
MergeDataActivity <- MergeDataActivity[order(MergeDataActivity$date), ]
```

## 7 - Histogram of the total number of steps taken each day after missing values are inputed.

```{r}
HistMergeDataActivity <- tapply(MergeDataActivity$steps, MergeDataActivity$date, sum)
hist(HistMergeDataActivity, xlab = "Number of Steps", main = "Histogram: Steps per Day (Imputed data)")
```

```{r}
MeanPerDayMergedDataActivity <- mean(HistMergeDataActivity, na.rm = TRUE)
MedianPerDayMergedDataActivity <- median(HistMergeDataActivity, na.rm = TRUE)
```

There are no difference between MEAN with missing data: `r MeanPerDayMergedDataActivity ` and replaced missing data: `r MedianPerDayMergedDataActivity` . but there is a little difference in Median: `r MeanPerDay` and `r MedianPerDay`.

## 8 - Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

## Creating new vector.

```{r}
MergeDataActivity$day <- ifelse(weekdays(as.Date(MergeDataActivity$date)) == "Saturday" | weekdays(as.Date(MergeDataActivity$date)) == "Sunday", "weekend", "weekday")
```
## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

# Average steps per interval for weekends
```{r}
StepsPerInterval.weekend <- tapply(MergeDataActivity[MergeDataActivity$day == "weekend" ,]$steps, MergeDataActivity[MergeDataActivity$day == "weekend" ,]$interval, mean, na.rm = TRUE)
```
# Average steps per interval for weekdays

```{r}
StepsPerInterval.weekday <- tapply(MergeDataActivity[MergeDataActivity$day == "weekday" ,]$steps, MergeDataActivity[MergeDataActivity$day == "weekday" ,]$interval, mean, na.rm = TRUE)
```
```{r}
par(mfrow=c(1,2))

plot(as.numeric(names(StepsPerInterval.weekday)), 
     StepsPerInterval.weekday, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekdays)", 
     type = "l")

plot(as.numeric(names(StepsPerInterval.weekend)), 
     StepsPerInterval.weekend, 
     xlab = "Interval", 
     ylab = "Steps", 
     main = "Activity Pattern (Weekends)", 
     type = "l")
```

